---
title: "Linear Models Week 1 Quiz"
author: "Nate Olson"
date: "2/13/2017"
output: html_document
---

```{r}
library(tidyverse)
```

1. 
The derivative (expressed as a column vector) with respect to X of AX+b, where A is an k×p matrix, X is a p×1 vector and b is a k×1 vector is? (Use a superscript t for transpose if needed).  

* X  

* $b$  

* Correct Answer: $A^t$  


2. 
The derivative (expressed as a column vector) with respect to X of (X−b)tA(X−b) where X and b are p×1 vectors and A is a p×p symmetric matrix is?  

* 2(X−b)  

* A  

* Correct Answer: $2A(X-b)$  

3. 
Let X be an n×p matrix. The matrix $(I-\frac{1}{n}J_{n,n})X(I−\frac{1}{p}J_{p,p})$ has all row and column means equal to zero.

* Correct Answer: True

* False

Verifying in R. Also in notes, formula used to calculate row and column means.

```{r}
X <- matrix(sample(1:100, size = 12),4,3)
n <- 4
In <- diag(n)
Jnn <- matrix(1,n,n)
p <- 3
Ip <- diag(p)
Jpp <- matrix(1,p,p)
```

```{r}
X
```

```{r}
In
```

```{r}
Jnn
```

```{r}
Ip
```

```{r}
Jpp
```

 $(I−1_nJ_{n,n})X(I−1_pJ_{p,p})$
```{r}
1/n * Jnn
```

```{r}
In - 1/n * Jnn
```

```{r}
Ip - 1/p * Jpp
```

```{r}
(In - 1/n * Jnn) %*% X
```

```{r}
X %*% (Ip - 1/p * Jpp)
```

```{r}
ans <- (In - 1/n * Jnn) %*% X %*% (Ip - 1/p * Jpp)
```

```{r}
colMeans(ans)
```
```{r}
rowMeans(ans)
```

4. 
Let X be an n×p matrix. The elements of $\frac{1}{n−1} X^t (I−\frac{1}{n} J_{n,n})X$ are what? ($J_{n,n}$ is an n×n matrix of ones.)
__Not sure why row and not columns, think has to do with ordering.__

* The empirical means of the columns of X.

* The empirical variances and covariances of the rows of X.

* Correct Answer: The empirical variances and covariances of the columns of X.

* The empirical means of the rows of X.

```{r}
t(X)
```

```{r}
1/(3-1) * t(X) %*% (In - 1/n*Jnn) %*% X
```

5. 
Let X be an n×p matrix with p×p variance/covariance matrix S. Let D be the p×p diagonal matrix of the square root of the diagonal entries of S. The matrix $XD^{−1}$ has what property?

```{r}
# X is a n x p matrix
X <- matrix(sample(1:100, 12), 4, 3)
S <- cov(X)
D <- S %>% diag() %>% diag() %>% sqrt()
x1 <- X %*% solve(D)
```


- All of its column means are 0.
```{r}
x1[, 1] %>% mean()
x1[, 2] %>% mean()
x1[, 3] %>% mean()
```

- All of its row means are 0.
```{r}
x1[1, ] %>% mean()
x1[2, ] %>% mean()
x1[3, ] %>% mean()
```


- All of its row variances are 1.
```{r}
x1[1, ] %>% var()
x1[2, ] %>% var()
x1[3, ] %>% var()
```


- __X__ All of its column variances are 1.
```{r}
x1[, 1] %>% var()
x1[, 2] %>% var()
x1[, 3] %>% var()
```


6. 
Let X be an n×p matrix with p×p variance/covariance matrix S. Let $\tilde{X} = (1−\frac{1}{n}J_{n,n})X$. Let D be the p×p diagonal matrix of the square root of the diagonal entries of S. The matrix $\frac{1}{n−1}D^{−1}\tilde{X}^t\tilde{X}D^{−1}$ has what properties?

```{r}
tildeX <- 1 - 1/(n-1) * Jnn %*% X
```

```{r}
x2 <- 1/(1/n) * solve(D) %*% t(tildeX) %*% tildeX %*% solve(D)
```

```{r}
x2
```

* Entries have the units of the columns of the X matrix squared.

* Entries have the units of the columns of the X matrix.

Entries are unitless


* Correct Answer: Has ones on its diagonals.

* Correct Answer: Has empirical correlations in the off diagonals

$\frac{1}{n - 1}D ^{ -1}\tilde{X}^t\tilde{X}D^{ -1} = D^{ -1} S D^{ -1}$. The purpose of this problem is to show how to convert a variance/ covariance matrix into a correlation matrix via matrix multiplication.