---
output:
  md_document:
    md_extensions: +tex_math_dollars
    pandoc_args: --webtex
    toc: yes
    toc_depth: 2
    variant: markdown_github
  html_document:
    toc: yes
---

```{r include=FALSE}
library(knitr)
opts_chunk$set(fig.path='img/',
               fig.width=1080/120,
               fig.height=1080/120,
               dpi=120, fig.retina=2)
library(purrr)
```

<h1>Linear Algebra Review XVIII SVD</h1>

- Nate Olson
- October 25, 2016

__Singular value decomposition__ - matrix factorization that can be applied to rectangular matricies.


__Singlular Value Decomposition__

$$A = U\Sigma V^T$$

- $\textbf{A}$ is a $m \times n$ matrix with rank $r$  
- __$\Sigma$__ is a $m \times n$ matrix that includes a diagonal matrix of size $r \times r$, where $r$ is the rank of the factorized matrix.  
  - the diagonal values in the matrix are the singular values in decreasing order  
- $\textbf{U}$__ - left singular values  
- $\textbf{V}$ - right singular values  

### Steps to performing SVD
1. Find an orthogonal diagonalization of __$A^{T}A$__ - eigenvalues of $A^TA$ and corresponding orthonormal eigenvectors.  
2. Set up __$V$__ and __$\Sigma$__ based on the ordered eigenvalues  
3. Construct __$U$__ - composed of the normalized vectors from __$Av_i$__ 


Good youtube video on SVD https://youtu.be/P5mlg91as1c 

### SVD in R
Matrix from example 3 in book

```{r}
(a <- matrix(c(4,11,14,8,7,-2),ncol=3, byrow = TRUE))
```

R has a command for calculating SVD
http://ugrad.stat.ubc.ca/R/library/base/html/svd.html
```{r}
svd(a,nv = 3)
```

#### Stepwise
1. Find an orthogonal diagonalization of __$A^{T}A$__ - eigenvalues of $A^TA$ and corresponding orthonormal eigenvectors.  

```{r}
(a_eigen <- eigen(t(a) %*% a))
```

2. Set up __$V$__ and __$\Sigma$__ based on the ordered eigenvalues  

```{r}
(sigma <- diag(sqrt(a_eigen$values)))
```


```{r}
(V <- a_eigen$vectors)

```

3. Construct __$U$__ - composed of the normalized vectors from __$\frac{1}{\sigma_i} Av_i$__ 
```{r}
e1 <- V[,1] %>% as.matrix()
e2 <- V[,2] %>% as.matrix()

u1 <- a %*% e1 * 1/sigma[1,1]
u2 <- a %*% e2 * 1/sigma[2,2]

(U <- cbind(u1,u2))
```

